---
title: "Gelman2012 HW1: Voting Data Example"
output: 
  html_notebook:
        toc: yes
        toc_float: true
---

# Loading packages

Our intention is to use packages from the tidyverse and other *tidy* packages to provide a constistant framework.      



```{r}

library(readr) # reading data files
library(lubridate) # working with dates
library(broom)
library(sf) #mapping and geospatial
library(tmap)
library(GISTools)
library(janitor) # data cleaning
library(tidyverse) # data manipulation and visualization - loads several packages
```

Some addition packages as indicated for the Bayesian work:


```{r}
library(foreign)
```


# Reading Data and preprocessing

Read the data.


```{r}
dffull=read.dta("pew_research_center_june_elect_wknd_data.dta")
df=read_csv("2008ElectionResult.csv") %>%  mutate(stateAbb=openintro::state2abbr(state))
```

Parse dataset to fields of interest to ease working with the data.

```{r}
# dfA=dffull %>% select(state,ideo) %>% group_by(state,ideo) %>% add_tally() %>% 
#     mutate(vl=p/n,stateAbb=openintro::state2abbr(state))

dfA=dffull %>% select(state,ideo) %>% tabyl(state,ideo) %>% adorn_totals("col") %>%
   mutate(vl=`very liberal`/Total,stateAbb=openintro::state2abbr(state))

dfA




```

We combine the polling data with the election results into a single dataframe.  We remove Alaska and Hawaii because we have too few polling results.

```{r}
dfB=dfA %>% select(-state) %>% left_join(df,by="stateAbb") %>%
  select(state,vl,everything()) %>% filter(Total > 1)
```

# Creating plots

The plot using the sampled value of percentage that are very liberal.

```{r}
dfB %>% ggplot(aes(x=vote_Obama_pct,y=vl,label=stateAbb)) +
  geom_label()+theme_minimal()+ggtitle('Survey percent of very liberal')
```

### Computing a prior distribution

Following the discussion from section 2.7 in the book, we treat each of the measured values of `vl` as being taken from a mixture of Poisson distributions.  As the book indicates, if that mixture is assumed to have come from poisson distributions whose rate parameters are Gamma distributed, then the result is a negative binomial distrubtion.   For our prior distribution for the rate parameter, we want to recover the Gamma distribution that would best represent the collection of rates across the states.

As described on page 51, we compute as follows:


```{r}
A=mean(dfB$vl)
B=var(dfB$vl)
C=mean(1/dfB$Total)
```

Then $A=\alpha / \beta,$

$B=CA +A/\beta .$

Solving for $\alpha$ and $\beta$ gives the values of the prior parameters:

```{r}
beta=A/(B-C*A)
alpha=A*beta
```

### Compute posterior distribution parameters

Following the method described on page 48 of the book, we computer poster parameters and mean for all states.
```{r}
dfB = dfB %>% mutate(postmean=(alpha+`very liberal`)/(beta+Total))

dfB %>% ggplot(aes(x=vote_Obama_pct,y=postmean,label=stateAbb)) +
  geom_label()+theme_minimal()+ggtitle('Posterior mean of very liberal')

```


### How much difference is there between the survey and the posterior

```{r}
dfB %>% ggplot(aes(x=vl,y=postmean,label=stateAbb)) +
  geom_label()+theme_minimal()+ggtitle('Posterior mean versus survey')

```





```{r}
dfB %>% ggplot(aes(x=Total,y=vl,label=stateAbb)) +
  geom_label()+theme_minimal()+ggtitle('Very Liberal versus number surveyed')

```

```{r}
dfB %>% ggplot(aes(x=Total,y=postmean,label=stateAbb)) +
  geom_label()+theme_minimal()+ggtitle('Posterior mean versus  number surveyed')

```