---
title: "Gelman Heirarchical Rats (Ch 5)"
output: html_notebook
---

## Book Example - section 5.3: Rats 

My intention here is to step through the procedures shown in the book for analysis of the rats data with heirarchical model.  

My general approach in working with R is biased by the type of work I generally do.  As such, I try to use dataframes as my primary data structures and uses commands from the tidyverse to maniuplate that data.  

>Note - dataframes and the tidyverse is *generally* a good approach, but not always.  Moreover, _there is more than one way to skin a cat._  So, your code would not need to look like mine.)


#### Packages

I load a standard kid of packages.  

```{r}
library(knitr)
library(lubridate) # working with dates
library(broom)
library(gridExtra)
library(tidyverse) # data manipulation and visualization - loads several packages

```

### Get the data

```{r}
df1=tibble(y = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,
                 1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,
        2,1,5,2,5,3,2,7,7,3,3,2,9,10,
        4,4,4,4,4,4,4,10,4,4,4,5,11,12,
        5,5,6,5,6,6,6,6,16,15,15,9,4),
n = c(20,20,20,20,20,20,20,19,19,19,19,
      18,18,17,20,20,20,20,19,19,18,18,25,24,
       23,20,20,20,20,20,20,10,49,19,46,27,
      17,49,47,20,20,13,48,50,20,20,20,20,
       20,20,20,48,19,19,19,22,46,49,20,
      20,23,19,22,20,20,20,52,46,47,24,14))
```

### Estimating $p(\alpha,\beta |y)$

Define our log-likelihood and posterior probability functions:

```{r}
logl <- function(a, b){
  lgamma(a+b)-lgamma(a)-lgamma(b)+lgamma(a+df1$y)+lgamma(b+df1$n-df1$y)-lgamma(a+b+df1$n)
}

fun=function (A,B) {-5/2*log(A+B)+(sum(logl(A,B)))}
```

To compute over an "appropriate" grid, we follow the book advice and sample on a grid in the  transformed coordinates.
The book give some help on reasonble ranges to build our grid

```{r}
Ax=seq(-2.3,-1.3,length.out=100) # values of ln(alpha/beta)
Bx=seq(1,5,length.out=100) # values of ln(alpha + beta)
```

We now build our data frame.  `Ax` and `Bx` represent the transformed variables, while `A` and `B` are the associated $\alpha$ and $\beta$ values.  I need those columns to evaluate the posterior function.

```{r}

df2=crossing(Ax,Bx) %>%
  mutate(A=exp(Ax+Bx)/(exp(Ax)+1),B=exp(Bx)/(exp(Ax)+1)) %>%
  mutate(p=pmap_dbl(list(A=A,B=B),fun),p=exp(p-max(p)))

```

### Visualize the posterior

```{r eval=FALSE, include=FALSE}
#xl <- c(-2, 8)
#yl <- c(-2, 40)

# df2 %>% ggplot(aes(x=A,y=B,)) +
#   geom_raster(aes(fill = p, alpha = p), interpolate = T) +
#   geom_contour(aes(z = p), colour = 'black', size = 0.2) +
#   labs(x = '~alpha', y = '~beta') +
#   scale_fill_gradient(low = 'yellow', high = 'red', guide = F) +
#   scale_alpha(range = c(0, 1), guide = F) +
#   theme_minimal()

```


```{r}
#xl <- c(-2, 8)
#yl <- c(-2, 40)

#df2 %>% ggplot(aes(x=log(A/B),y=log(A+B))) +
 # geom_point()
  
 df2 %>% ggplot(aes(x=Ax,y=Bx,)) +
  geom_raster(aes(fill = p, alpha = p), interpolate = T) +
  geom_contour(aes(z = p), colour = 'black', size = 0.2) +
  labs(x = bquote(ln(alpha/beta)), y = bquote(ln(alpha+beta))) +
  scale_fill_gradient(low = 'yellow', high = 'red', guide = F) +
  scale_alpha(range = c(0, 1), guide = F) +
  theme_minimal()
```

We may sample from that posterior and visualize:

```{r}
nsamp=1000
mysample=df2 %>% sample_n(size=nsamp, replace=TRUE,weight=p)
```

in the space of the transformed coordinates.
```{r}
mysample %>% ggplot( aes(x=Ax, y=Bx))    +
  geom_jitter(color = 'blue') +
  labs(x = bquote(ln(alpha/beta)), y = bquote(ln(alpha+beta))) +
  theme_minimal()
```

And in the original hyperparameter space.

```{r}
mysample %>% ggplot( aes(x=A, y=B))    +
  geom_jitter(color = 'blue') +
  labs(x = 'alpha', y = 'beta') +
  theme_minimal()
```


### Simulations

Let's now consider the simulation of the full model (as outlined in section 5.3).  

#### Step 1
`mysample` already gives us our 1000 draws from posterior marginal $p(\alpha,\beta|y)$.

#### Step 2

For each of those draws, we want to get our `J=71` values for $\theta_j,$ drawn from the posterior conditional distribution $p(\theta_j|\alpha,\beta,y)$ taken from distribution $Beta(\alpha+y_j, \beta+n_j-y_j)$.

```{r}
a1=mean(mysample$A)
b1=mean(mysample$B)

dfs=mysample %>% select(A,B,p)
dfs$theta71.post=rbeta(1,dfs$A+4,dfs$B+14)
dfs$theta71.post.res=rbeta(1,a1+4,b1+14)

#dfs2=df2 %>% gather

```


```{r}
dfs %>% ggplot(aes(x=theta71.post))+geom_density(aes(x=theta71.post),color="Red")+
  geom_density(aes(x=theta71.post.res),color="Blue")+
  xlim(0, .6)+
theme_minimal()
```






