---
title: "Stan analysis of ABC1 Testing results using Heirarchical model"
author: "Joe Skufca"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

I will use the stan code from the Stan User's Guide.



### Packages

I load a standard kid of packages, in case we need it.

```{r}
library(knitr)
library(broom)
library(gridExtra)
library(readxl)
library(tidyverse) # data manipulation and visualization - loads several packages
library(janitor)
```
And I also load `rstan` an R wrapper that makes it easy to use STAN inside of R.



```{r}
library ("rstan")
rstan_options(auto_write = TRUE)
```


### Load a dataset

From 2018, data is test scores for first 4 exams, filtered so that we are looking only at students who took all four exams (to simplify our structure).

```{r}
df1=read_excel("MA131_ABC_EXAM_Fall_2018_final DL.xls",col_names = F) %>%
  rename_with(~paste0("t",.x))%>% clean_names() %>% 
  mutate(ID=factor(1:139))

df2=df1 %>% pivot_longer(1:4) %>% select(-name) %>% mutate(t=rep(c(0,1,2,3),139))

df2 %>% group_by(ID) %>%  ggplot(aes(x=t,y=value,group=ID)) +geom_line()

```


### Form our data

```{r}
N=139*3;
J=139;

df3=df1 %>%  pivot_longer(1:3) %>% select(-name) %>% mutate(t=rep(c(0,1,2),139))
id=as.integer(df3$ID) 
X=df3$t
y=df3$value


```

### Fit the model

```{r}
reg_fit=stan(file = "regressionABC.stan",
                 data=c("N","J","id","X","y"), iter=1000,chains = 4)
print(reg_fit)
plot(reg_fit)

```

# Below not fixed

### Run the stan using a chaing





```{r}
plot(reg_fit,plotfun="rhat",ncol=1)
```

```{r}

```


