---
title: "Stan example: Estimating a location"
output:
  html_notebook:
    theme: united
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
subtitle: Motivated by Tianqi's research
editor_options:
  chunk_output_type: inline
---

I am working on the problem of estimating the location of an object based purely on the set of distances from some fixed anchor locations.

## Packages

I load a standard kid of packages, in case we need it.

```{r}
library(knitr)
library(lubridate) # working with dates
library(broom)
library(gridExtra)
library(tidyverse) # data manipulation and visualization - loads several packages

```
And I also load `rstan` an R wrapper that makes it easy to use STAN inside of R.

```{r message=FALSE, warning=FALSE}
library ("rstan")
library("loo")
rstan_options(auto_write = TRUE) # recommended option - writes the compiled model
options(mc.cores = parallel::detectCores())
```


## Problem 1: a single object

For this first model, let's assume we are trying to locate a single object by observing is distance from N=5 anchors (fixed sensors).   

### Data

We first create surrogate data.  The anchors are resumed to be at locations defined in $X$, where each row is an anchor position.  We will assume a known location of $(0.6,0.8)$ for the object and create surrogate data for the location information by using a lognormal distrution (representing muliplicative error)

```{r}
N=5 # number of anchors
X=matrix(c(0,.5,1,1.5,2,0,3,0,1,4),N,2)
plot(X,asp=1)

d=sqrt((X[,1]-.6)^2+(X[,2]-.8)^2) # true distances to (0.6,0.8)

meas.sd=.1
y=rlnorm(5,log(d),meas.sd)
y
d

```

### Create the stan model

Good work practices are to create the model in a `.stan` file, which allows for easier reuse, to include using that file with other interfaces that access the stan library.  We will create in a file called `TQdist1.stan`.

Some notes in building the model:

* You can access the stan User guide and reference manual at https://github.com/stan-dev/stan/releases/download/v2.17.0/stan-reference-2.17.0.pdf
* For Gelman's advice on priors in STAN, take a look at https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
* If you understand the *Tidy Data* construct in R, then applying to R models makes use of the *tidybayes* package.  See a short tutorial at https://mjskay.github.io/tidybayes/articles/tidybayes.html





### Fit the stan model

The `stan` command takes the *.stan* model file and compiles the required C-code and saves it.  (Turns out, this can take a while.)  Then it generates the necessary chains and samples.  If you run it a second time, it has already compiled the code, and will only do the sampling.

```{r}
TQ1_fit=stan(file = "TQdist1.stan", 
              data=c("N","X","y"),
             iter=1000,chains = 4)
print(TQ1_fit)
plot(X,pch=2,col="blue",asp=1)
points(extract(TQ1_fit)$theta)


 
```

If we limit the range of possible locations.

```{r}
TQ2_fit=stan(file = "TQdist2.stan", 
              data=c("N","X","y"),
             iter=1000,chains = 4)
print(TQ2_fit)
plot(X,pch=2,col="blue",asp=1)
points(extract(TQ2_fit)$theta,asp=1)


```


If we specifiy a constraint that $theta_1+theta_2=1.4$.

```{r eval=FALSE, include=FALSE}
TQ3_fit=stan(file = "TQdist3.stan",
              data=c("N","X","y"),
             iter=1000,chains = 4)
print(TQ3_fit)
plot(X,pch=2,col="blue",asp=1)
points(extract(TQ3_fit)$theta)

```

Now lets create a model as if we expected the position to be on a road with some width.

```{r}
TQ4_fit=stan(file = "TQdist4.stan",
              data=c("N","X","y"),
             iter=1000,chains = 4)
print(TQ4_fit)
plot(X,pch=2,col="blue",asp=1)
points(extract(TQ4_fit)$theta)
```


Can I build a heirarchical model with many agents on a road - and discover the road.


