---
title: "Stan example of simple linear regression"
author: "Joe Skufca"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

I will use the stan code from the Stan User's Guide.

https://mc-stan.org/docs/2_24/stan-users-guide/linear-regression.html



#### Packages

I load a standard kid of packages, in case we need it.

```{r}
library(knitr)
library(broom)
library(gridExtra)
library(tidyverse) # data manipulation and visualization - loads several packages

```
And I also load `rstan` an R wrapper that makes it easy to use STAN inside of R.



```{r}
library ("rstan")
rstan_options(auto_write = TRUE)
```


### Create a dataset

```{r}
df1=cars


N = 50
x=df1$speed
y=df1$dist


```

### Run the stan using a chaing


```{r}
reg_fit=stan(file = "regression1.stan",
                 data=c("N","x","y"), iter=1000,chains = 4)
print(reg_fit)
plot(reg_fit)

```


```{r}
plot(reg_fit,plotfun="rhat",ncol=1)
```

```{r}
lm(dist~speed,data=df1) %>% summary()
```

## A second example - a learning problem

Suppose we "learn" a task (over time), with $p$ giving the probability that we can demonstrate a particular knowledge element.

### Data creation

```{r}
A=.95
B=.15

N = 20
t=1:N # data

p=A*(1-exp(-B*t)) # unknown 

y=rbinom(20,10,p) # data

```


```{r}
reg_fit=stan(file = "regression2.stan",
                 data=c("N","t","y"), iter=1000,chains = 4)
print(reg_fit)
plot(reg_fit)

```



