---
title: "Stan example from Appendix C"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

I am working from page 590 in the book, to show how stan would be used on a problem we have seen before.

#### Packages

I load a standard kid of packages, in case we need it.

```{r}
library(knitr)
library(lubridate) # working with dates
library(broom)
library(gridExtra)
library(tidyverse) # data manipulation and visualization - loads several packages

```
And I also load `rstan` an R wrapper that makes it easy to use STAN inside of R.



```{r}
library ("rstan")
rstan_options(auto_write = TRUE)
```


### Get the data

```{r}
J <- 8
y <- c(28,8,-3,7,-1,1,18,12)
sigma <- c(15,10,16,11,9,11,10,18)
```

### Run the stan using a chaing


```{r}
schools_fit=stan(file = "schools.stan",
                 data=c("J","y","sigma"), iter=1000,chains = 4)
print(schools_fit)
plot(schools_fit)

```

## Let's try the Rats example

We now consider a separate example looking at grow rate of rats.  Data taken from https://raw.github.com/wiki/stan-dev/rstan/rats.txt .


```{r}
y <- as.matrix(read.table('https://raw.github.com/wiki/stan-dev/rstan/rats.txt', header = TRUE))
x <- c(8, 15, 22, 29, 36)
xbar <- mean(x)
N <- nrow(y)
T <- ncol(y)
```


```{r}
#rats_fit=stan(file = "rats.stan",iter=1000,chains = 4)
tictoc::tic("rats")
rats_fit <- stan('rats2.stan')
tictoc::toc()
print(rats_fit)
plot(rats_fit)

```


